---
title: "Analysis of Bryophyte Richness and Diversity in the Americas"
author: Carter Powell, Jackie O'Malley, and Julia Eckberg
output: html_notebook
---

```{r}
require(knitr)
require(BIEN)
require(ape)
require(maps)
require(dplyr)
require(maptools)
require(raster)
require(dismo)
require(sp)
require(rgdal)
require(mapdata)
require(mapproj)
require(vegan)
require(letsR)
require(entropart)
require(betapart)
require(rgeos)
require(CommEcol)
require(rlist)
```

#Load and process data
```{r, eval=TRUE, echo=FALSE}
source("~/Documents/Bryophytes/Scripts/DataLoading.R")
source("~/Documents/Bryophytes/Scripts/DataProcessing.R")
```


#Create occurrence by cell matrix by reshaping dataframe
```{r}
require(reshape2)

SpeciesCellID <- BryophytePresence[,c(1,4)]
melted <- melt(SpeciesCellID, id=c("Species", "CellID"), na.rm = TRUE)

SpeciesCellMatrix <- acast(melted, CellID~Species, margins=TRUE)
```


#Using betadiver to compute B-diversity using Jaccard dissimilarity
```{r}
betaj <- betadiver(SpeciesCellMatrix, method = "j", order = FALSE, help = FALSE)
head(betaj)
#betavector <- as.vector(t(betaj))
```


#Using vegdist to compute B-diversity using presence-absence data
```{r}
#betapresab <-vegdist(SpeciesCellMatrix, method="jaccard", binary=TRUE, diag=FALSE, upper=TRUE, na.rm = FALSE)
```


#Identify cells that are adjacent
```{r}
cellID <- tally$CellID

cellvector <- c(1:15038)
neighbor <- function(cellvector) {(adjacent(BIENblank, cellvector, directions=8, pairs=FALSE, target=NULL, sorted=TRUE, include=FALSE, id=FALSE))}
neighbors <- lapply(cellvector, neighbor)
names(neighbors) <- cellvector

bryneighbors <- neighbors[cellID]
```

#Separate out cells with 8 neighbors
```{r}
neighborvect <- unlist(lapply(bryneighbors, length))

cell8 <- cellID[which(neighborvect==8)]

betaj<-as.matrix(betaj)
betaj[!cell8, !cell8, drop=TRUE]

inx <- match(as.character(cell8), rownames(betaj))

betaj <- betaj[inx,inx]

neighbormatrix <- matrix(as.data.frame(cell8))
```

#Trying to extract the vector of neighbors for each cell with 8 neighbors
```{r}
#neighborvect <- unlist(lapply(bryneighbors))
#mergedlist<-cbind(bryneighbors, as.vector(neighborlist))

neighborlist <- as.list(cell8)
bry8neighbors <- neighbors[cell8]
```

#Try out beta raster
```{r}
BetaRaster <- setValues(BIENblank, betavector)
```

#Trying out betadisper
```{r}
betadisper(betaj, group, type = c("median","centroid"), bias.adjust = FALSE,
       sqrt.dist = FALSE, add = FALSE)
```

#Turn occurrence matrix into presence-absence matrix
```{r}
#SpeciesCellMatrix[SpeciesCellMatrix > 0] <- 1
```

#Compute Jaccard dissimilarity for all pairs of sites
```{r}
#JaccardMatrix <- vegdist(SpeciesCellMatrix, method="jaccard", binary=TRUE, upper=TRUE)
```

#Convert blank raster to shapefile
```{r, eval=FALSE}
crs_ref <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"  
BIENblank_tmp <- projectRaster(BIENblank, crs = crs_ref)

BIENpoints <- rasterToPoints(BIENblank_tmp, spatial  = TRUE)
proj4string(BIENpoints)

BIENpoints <- spTransform(BIENpoints, CRS(crs_ref))

BIENpoints@data <- data.frame(BIENpoints@data, long=coordinates(BIENpoints) [,1], lat=coordinates(BIENpoints) [,2])
head(BIENpoints@data, n=20)

BIENpolygons <- rasterToPolygons(BIENblank_tmp, fun = NULL, na.rm = TRUE, n=8, digits = 12)
BIENpolygons
```


#Trying out betagrid
```{r}
#print(betagrid)
#betagrid(gridshp=BIENpolygons, comp = SpeciesCellMatrix, xfeature = 2, yfeature = 3, radius = 0.25,index = "jaccard")
```


#Trying out betagrid2
```{r}
#PresAbTable <- read.table(SpeciesCellMatrix, row.names=1, head=T)
#print(betagrid2)
#betagrid2(SpeciesCellMatrix, radius = 100)
```
#Turn occurrence matrix into presence-absence matrix
