---
title: "Analysis of Bryophyte Richness and Diversity in the Americas"
author: Carter Powell, Jackie O'Malley, and Julia Eckberg
output: html_notebook
---

```{r}
require(knitr)
require(latexpdf)
require(sp)
require(vegan)
require(raster)
require(rasterVis)
require(wesanderson)
require(ggplot2)
require(gridExtra)
require(sf)
```


#Load blank raster and cell richness data + extract cell IDs and create vector for all cells
#Change file for BetaMat and CellRichness depending on if you want to map bryophytes, mosses, liverworts, etc. 

```{r}
BlankRas <-raster("Data/blank_100km_raster.tif")
BetaMat <- readRDS("Data/BetaMat.rds")
CellRichness <- readRDS("Data/CellRichness.rds")
CellID <- CellRichness$CellID
CellVec <- c(1:15038)
```


#Identify occupied cells that are adjacent to each occuppied cell + convert to vector
```{r}
neighbor <- function(CellVec) {(adjacent(BlankRas, CellVec, directions=8, pairs=FALSE, target=CellID, sorted=TRUE, include=FALSE, id=FALSE))}
Neighbors <- lapply(CellVec, neighbor)
names(Neighbors) <- CellVec

bryneighbors <- Neighbors[CellID]
bryneighborvect <- unlist(lapply(bryneighbors, length))
```


#Separate out occuppied cells with 8 and 7 occuppied neighbors
```{r}
Cell8 <- CellID[which(bryneighborvect==8)]
Neighbors8 <-Neighbors[Cell8]
Neighbors8 <- data.frame(Neighbors8)
names(Neighbors8) <- Cell8

Cell7 <- CellID[which(bryneighborvect==7)]
Neighbors7 <- Neighbors[Cell7]
Neighbors7 <- data.frame(Neighbors7)
names(Neighbors7) <- Cell7
```


#Make beta diversity matrix for all cells
```{r}
BetaMat<-as.matrix(BetaMat)
row.names(BetaMat) <- CellID
names(BetaMat) <- CellID
```


#Make beta diversity matrix for cells with 8 neighbors and cells with 7 neighbors
```{r}
BetaMat8<- BetaMat[!Cell8, !Cell8, drop=TRUE]
inx8 <- match(as.character(Cell8), rownames(BetaMat8))
BetaMat8 <- BetaMat8[inx8,inx8]

BetaMat7 <- BetaMat[!Cell7, !Cell7, drop=TRUE]
inx7 <- match(as.character(Cell7), rownames(BetaMat7))
BetaMat7 <- BetaMat7[inx7,inx7]
```


#For each cell, pairwise beta diversity is calculated for that focal cell and each of its 8 (or 7) neighbors, and the mean of those values is found
```{r}
Cell8CH <- as.character(Cell8)
Beta8 <- lapply(Cell8CH, function(x)mean(BetaMat[x, as.character(Neighbors8[,x])]))
names(Beta8) <- Cell8CH

Cell7CH <- as.character(Cell7)
Beta7 <- lapply(Cell7CH, function(x)mean(BetaMat[x, as.character(Neighbors7[,x])]))
names(Beta7) <- Cell7CH
```


#Plot mean pairwise beta diversity by Cell ID
```{r}
Beta7Vec<-unlist(Beta7)
Beta8Vec<-unlist(Beta8)

BetaVec <- rep(0, 15038)

BetaVec[Cell8]<-Beta8Vec
BetaVec[Cell7]<-Beta7Vec
BetaVec[BetaVec==0]<-NA

plot(BetaVec, ylab = "Mean Pairwise β-Diversity", xlab = "Cell ID")
```


#Load mapping packages and create colorscheme
```{r}
cols <- rev(wes_palette("Zissou1", 500, type = "continuous"))
```


#Add continental and mountainous outlines
```{r, message=FALSE}
nw_mount <- shapefile("Data/MapOutlines/Mountains/Koeppen-Geiger_biomes.shp")
nw_bound <- shapefile("Data/MapOutlines/Global_bound/Koeppen-Geiger_biomes.shp")

nw_mount_sf <- st_as_sf(nw_mount)
nw_bound_sf <- st_as_sf(nw_bound)
```


#Map beta diversity, adjust minimum value to remove outliers
```{r}
BetaRaster <- setValues(BlankRas, BetaVec)
BetaPoints<-rasterToPoints(BetaRaster)

BetaDF <- data.frame(BetaPoints)
colnames(BetaDF) <- c("Longitude", "Latitude", "Beta")

theme_set(theme_void())
RawBetaMap <- ggplot() + geom_tile(data=BetaDF, aes(x=Longitude, y=Latitude, fill=Beta)) + 
    scale_fill_gradientn(name="β-diversity", colours=cols, na.value="transparent", limits = c(0,1)) + 
    coord_equal()

RawBetaMap
```

#Subset beta values above 0.5
```{r}
HighBetaVec <- rep(0, 15038)
HighBetaVec[Cell8]<-Beta8Vec
HighBetaVec[Cell7]<-Beta7Vec
HighBetaVec[HighBetaVec==0]<-NA
HighBetaVec[HighBetaVec<0.5]<-NA

HighBetaRaster <- setValues(BlankRas, HighBetaVec)
HighBetaPoints<-rasterToPoints(HighBetaRaster)

HighBetaDF <- data.frame(HighBetaPoints)
colnames(HighBetaDF) <- c("Longitude", "Latitude", "Beta")
```


#Beta diversity by latitude
```{r}
HighBetaDF$pc <- predict(prcomp(~Latitude+Beta, HighBetaDF))[,1]
cols2 <- rev(wes_palette("Zissou1", 4, type = "continuous"))

theme_set(theme_gray())
BetaScatterplot <- ggplot(HighBetaDF, aes(Latitude, Beta, color=pc)) + geom_point(shape = 16, size = 5, show.legend = FALSE, alpha=0.5) + scale_color_gradientn(colours=cols2) + theme_minimal()

BetaScatterplot
```


#Map low beta values
```{r}
LowBetaVec <- rep(0, 15038)
LowBetaVec[Cell8]<-Beta8Vec
LowBetaVec[Cell7]<-Beta7Vec
LowBetaVec[LowBetaVec==0]<-NA
LowBetaVec[LowBetaVec>0.5]<-NA

LowBetaRaster <- setValues(BlankRas, LowBetaVec)
LowBetaPoints<-rasterToPoints(LowBetaRaster)
LowBetaDF <- data.frame(LowBetaPoints)
colnames(LowBetaDF) <- c("Longitude", "Latitude", "Beta")

theme_set(theme_void())
LowBetaMap <- gplot(LowBetaRaster, maxpixels=15038) +  geom_tile(aes(fill = value)) + 
    scale_fill_gradientn(colours="black", na.value="transparent", limits = c(0,1)) + theme(legend.position = "none") + 
    coord_equal() 
LowBetaMap
```


#Map beta diversity with values under 0.5 shown in dark grey (19 values are under 0.5)
```{r}
source("Functions/gplot_data.R")
gplotB<- gplot_data(BetaRaster)
gplotLB<- gplot_data(LowBetaRaster)


BetaMap <- ggplot() +
  geom_tile(data = dplyr::filter(gplotLB, !is.na(value)), 
            aes(x = x, y = y), fill = "gray25") +
  geom_tile(data = gplotB, 
            aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(name="β-diversity", colours=cols, na.value="transparent", limits = c(0.5,1)) +
  coord_quickmap() + geom_sf(data = nw_bound_sf, size = 0.1, fill=NA) + 
  geom_sf(data = nw_mount_sf, size = 0.2, alpha=0.1)
BetaMap
```


#Map richness (alpha diversity)
```{r, warning=FALSE}
RichnessVec <- readRDS("Data/RichnessVec.rds")
RichnessVec[which(RichnessVec==0)]=NA
RichnessRaster <- setValues(BlankRas, RichnessVec)

RichnessDF <- rasterToPoints(RichnessRaster)
RichnessDF <- data.frame(RichnessDF)

colnames(RichnessDF) <- c("Longitude", "Latitude", "Alpha")

theme_set(theme_void())
RichnessMap <- ggplot() + geom_tile(data=RichnessDF, aes(x=Longitude, y=Latitude, fill=Alpha)) +   
    scale_fill_gradientn(name="α-diversity", colours=cols, na.value="transparent") +
    coord_equal() +
 geom_sf(data = nw_bound_sf, size = 0.1, fill=NA) + 
  geom_sf(data = nw_mount_sf, size = 0.2, alpha=0.1)
```


#Alpha diversity by latitude
```{r}
RichnessDF$pc <- predict(prcomp(~Latitude+Alpha, RichnessDF))[,1]
cols2 <- rev(wes_palette("Zissou1", 4, type = "continuous"))

theme_set(theme_gray())
AlphaScatterplot <- ggplot(RichnessDF, aes(Latitude, Alpha, color=pc)) + geom_point(shape = 16, size = 5, show.legend = FALSE, alpha=0.5) + scale_color_gradientn(colours=cols2) + theme_minimal()

AlphaScatterplot

png("AlphaBetaScatter.png", width = 1000, height = 1000, pointsize = 20)
grid.arrange(AlphaScatterplot, BetaScatterplot, ncol=1)
dev.off()
```


#Map alpha and beta diversity side-by-side and save as png
```{r}
png("bothmaps.png", width = 1000, height = 1000, pointsize = 20)
grid.arrange(RichnessMap, BetaMap, ncol=2)
dev.off()
```


#Plot Jaccard index against Sorensen (both are SIMILARITY)
```{r}
SorBetaVec <- readRDS("Data/SorBetaVec.rds")
SorBetaVec[SorBetaVec<0]<-NA
BetaVec[BetaVec<0]<-NA

plot(BetaVec, ylab = "β-diversity", xlab = "Cell ID", col="darkgoldenrod2")
points(SorBetaVec, col="cadetblue4")
legend("bottomright", legend=c("Jaccard", "Sorensen"),
        fill=c("darkgoldenrod2","cadetblue4"), horiz=FALSE, cex=0.9)
```
