---
title: "Subsetting Mountains and Lowlands"
author: "Jackie O'Malley '21 and Julia Eckberg '21"
date: "7/25/2019"
output: html_document
---

```{r, message = FALSE}
require(sp)
require(raster)
require(ggplot2)
require(gridExtra)
require(rgdal)
```


#Add continental and mountainous region outlines
```{r, message = FALSE}
nw_mount <- shapefile("Data/MapOutlines/Mountains/Koeppen-Geiger_biomes.shp")
nw_bound <- shapefile("Data/MapOutlines/Global_bound/Koeppen-Geiger_biomes.shp")

LongLatBetaRaster <- readRDS("Data/LongLatBetaRaster.rds")
RichnessRaster <- readRDS("Data/RichnessRaster.rds")
BlankRas <-raster("Data/blank_100km_raster.tif")
CellVec <- c(1:15038)
```


#Creates dataframe including coordinates for each cell
```{r}
LongLatRaster <- setValues(BlankRas, CellVec)
LongLatPoints<-rasterToPoints(LongLatRaster)
LongLatDF <- data.frame(LongLatPoints)
colnames(LongLatDF) <- c("Longitude", "Latitude", "CellID")

coordinates(LongLatDF) <- ~Longitude+Latitude 
proj4string(LongLatDF) <- CRS("+proj=utm +zone=10") 
LongLat <- spTransform(LongLatDF, CRS("+proj=longlat")) 
LongLatDF <- data.frame(LongLat)
LongLatDF[c("Longitude", "Latitude", "CellID")]
LongLatDF <- subset(LongLatDF, select = -c(optional))
saveRDS(LongLatDF, "Data/LongLatDF.rds")

```

#Binning
```{r}
#LongLatDF$Bin = cut(LongLatDF$Latitude,c(-70,-30,0,30,70), labels = c("M1, M2, M3, M4"))
#table(LongLatDF$Bin)
```


#Scatterplot of beta diversity values of cells whose centers are within mountainous regions 
```{r, warning = FALSE}
LongLatDF <- readRDS("Data/LongLatDF.rds")
BetaMount <- extract(LongLatBetaRaster, nw_mount, df = TRUE, cellnumbers = TRUE)

colnames(BetaMount) <- c("Type", "CellID", "Beta")
BetaMount$Type <- "Mountain"

BetaMountVec <- BetaMount$CellID
BetaMount <- merge(BetaMount, LongLatDF)

BetaMountScatterplot <- ggplot() + geom_point(data = BetaMount, aes(Latitude, Beta), shape = 16, size = 5, show.legend = FALSE, alpha=0.5, color = "goldenrod2") + ylab("Mountainous β diversity") + ylim(0.3, 1.0) + xlab("Latitude") + theme_minimal() +  theme(axis.title.y = element_text(size=32), axis.title.x = element_text(size=32),  axis.text = element_text(size=20))
BetaMountScatterplot
```

#Bin mountains by 10 degrees
```{r}
BetaMount$Bin10 = cut(BetaMount$Latitude,c(-70,-60,-50,-40,-30,-20,-10,0,10,20,30,40,50,60), labels = c("1M", "2M", "3M", "4M", "5M", "6M", "7M", "8M", "9M", "10M", "11M", "12M", "13M"))
table(BetaMount$Bin10)
```

#Bin mountains by 5 degrees
```{r}
BetaMount$Bin5 = cut(BetaMount$Latitude,c(-70,-65,-60,-55,-50,-45,-40,-35,-30,-25,-20,-15,-10,-5,0,5,10,15,20,25,30,35,40,45,50,55), labels = c("1M", "2M", "3M", "4M", "5M", "6M", "7M", "8M", "9M", "10M", "11M", "12M", "13M", "14M", "15M", "16M", "17M", "18M", "19M", "20M", "21M", "22M", "23M", "24M", "25M"))
table(BetaMount$Bin5)
```

#Scatterplot of beta diversity values of cells in lowland areas 
```{r, warning = FALSE}
LongLatDF <- readRDS("Data/LongLatDF.rds")

BetaBound <- extract(LongLatBetaRaster, nw_bound, df = TRUE, cellnumbers = TRUE)
colnames(BetaBound) <- c("Type", "CellID", "Beta")

BetaMatch <- BetaBound[BetaBound$CellID %in% BetaMountVec, ]
BetaMatchVec <- BetaMatch[, "CellID"]
BetaLowland <- BetaBound[!BetaBound$CellID %in% BetaMatchVec, ]

BetaLowland$Type <- "Lowland"

BetaLowlandVec <- BetaLowland$CellID
BetaLowland <- merge(BetaLowland, LongLatDF)

BetaLowlandScatterplot <- ggplot() + geom_point(data = BetaLowland, aes(Latitude, Beta), shape = 16, size = 5, show.legend = FALSE, alpha=0.5, color = "cyan4") + ylab("Lowland β diversity") + ylim(0.3, 1.0) + xlab("Latitude") + theme_minimal() + theme(axis.title.y = element_text(size=32), axis.title.x = element_text(size=32),  axis.text = element_text(size=20))
BetaLowlandScatterplot

png("Figures/BetaMountainLowlandScatter.png", width = 1500, height = 1000, pointsize = 20)
grid.arrange(BetaMountScatterplot, BetaLowlandScatterplot, ncol=1)
dev.off()
```

#Bin lowlands by 10 degrees
```{r}
BetaLowland$Bin10 = cut(BetaLowland$Latitude,c(-70,-60,-50,-40,-30,-20,-10,0,10,20,30,40,50,60), labels = c("1L", "2L", "3L", "4L", "5L", "6L", "7L", "8L", "9L", "10L", "11L", "12L", "13L"))
table(BetaLowland$Bin10)
```

#Bin lowlands by 5 degrees
```{r}
BetaLowland$Bin5 = cut(BetaLowland$Latitude,c(-65,-60,-55,-50,-45,-40,-35,-30,-25,-20,-15,-10,-5,0,5,10,15,20,25,30,35,40,45,50,55,60), labels = c("2L", "3L", "4L", "5L", "6L", "7L", "8L", "9L", "10L", "11L", "12L", "13L", "14L", "15L", "16L", "17L", "18L", "19L", "20L", "21L", "22L", "23L", "24L", "25L", "26L"))
table(BetaLowland$Bin5)
```


#BoxPlots
```{r}
require(forcats)
FullBeta <- rbind(BetaMount, BetaLowland)

reordered10 <- fct_relevel(FullBeta$Bin10, "1L", "1M","2L", "2M", "3L", "3M", "4L", "4M", "5L", "5M", "6L", "6M", "7L", "7M", "8L", "8M","9L", "9M", "10L", "10M", "11L", "11M", "12L", "12M", "13L", "13M")

reordered5 <- fct_relevel(FullBeta$Bin5, "1M","2L", "2M", "3L", "3M", "4L", "4M", "5L", "5M", "6L", "6M", "7L", "7M", "8L", "8M","9L", "9M", "10L", "10M", "11L", "11M", "12L", "12M", "13L", "13M", "14L", "14M","15L", "15M", "16L", "16M", "17L", "17M", "18L", "18M", "19L", "19M", "20L", "20M", "21L", "21M","22L", "22M", "23L", "23M", "24L", "24M", "25L", "25M", "26L")

BetaBoxPlot <- ggplot(data = FullBeta, aes(reordered10, Beta, fill=Type)) + 
  geom_boxplot() + 
  ylab("β diversity") + ylim(0.5, 1.0) + xlab("Latitude") + 
  theme_minimal() + theme(axis.title.y = element_text(size=40), axis.title.x = element_text(size=40),  
                          axis.text.x = element_blank(), axis.text.y = element_text(size=20), legend.text = element_text(size=32), 
                          legend.position = "bottom", legend.title = element_blank()) + 
  scale_fill_manual(values = c("cyan4", "goldenrod2"))
BetaBoxPlot

png("Figures/BetaBoxPlot.png", width = 2000, height = 1000, pointsize = 20)
BetaBoxPlot
dev.off()
```


#Beta scatterplots combined
```{r, warning = FALSE}
FullBeta <- rbind(BetaMount, BetaLowland)

BetaScatter <- ggplot(data = FullBeta, aes(Latitude, Beta, color=Type)) + 
  geom_point(shape = 16, size = 3, alpha=0.8) + 
  ylab("β diversity") + ylim(0.5, 1.0) + xlab("Latitude") + 
  theme_minimal() + theme(axis.title.y = element_text(size=40), axis.title.x = element_text(size=40),  
                          axis.text = element_text(size=20), legend.text = element_text(size=32), 
                          legend.position = "bottom", legend.title = element_blank()) + 
  scale_color_manual(values = c("cyan4", "goldenrod2")) + geom_smooth(size = 2, show.legend = FALSE)
BetaScatter
```


#Scatterplot of alpha diversity values of cells whose centers are within mountainous regions 
```{r, warning = FALSE}
LongLatDF <- readRDS("Data/LongLatDF.rds")

AlphaMount <- extract(RichnessRaster, nw_mount, df = TRUE, cellnumbers = TRUE)

colnames(AlphaMount) <- c("Type", "CellID", "Alpha")
AlphaMount$Type <- "Mountain"

AlphaBoundVec <- AlphaMount[, "CellID"]
AlphaMount <- merge(AlphaMount, LongLatDF)

AlphaMountScatterplot <- ggplot(AlphaMount, aes(Latitude, Alpha)) + geom_point(shape = 16, size = 5, show.legend = FALSE, alpha=0.5, color = "goldenrod2") + ylab("Mountainous α diversity") + ylim (0, 800) + xlab("Latitude") + theme_minimal() + theme(axis.title.y = element_text(size=32), axis.title.x = element_text(size=32),  axis.text = element_text(size=20))
AlphaMountScatterplot

png("Figures/AlphaBetaMountScatter.png", width = 1000, height = 1000, pointsize = 20)
grid.arrange(AlphaMountScatterplot, BetaMountScatterplot, ncol = 1)
dev.off()
```


#Bin mountains by 5 degrees
```{r}
AlphaMount$Bin5 = cut(AlphaMount$Latitude,c(-70,-65,-60,-55,-50,-45,-40,-35,-30,-25,-20,-15,-10,-5,0,5,10,15,20,25,30,35,40,45,50,55), labels = c("1M", "2M", "3M", "4M", "5M", "6M", "7M", "8M", "9M", "10M", "11M", "12M", "13M", "14M", "15M", "16M", "17M", "18M", "19M", "20M", "21M", "22M", "23M", "24M", "25M"))
table(AlphaMount$Bin5)
```



#Scatterplot of alpha diversity values of cells in lowland areas 
```{r, warning = FALSE}
LongLatDF <- readRDS("Data/LongLatDF.rds")

AlphaBound <- extract(RichnessRaster, nw_bound, df = TRUE, cellnumbers = TRUE)
colnames(AlphaBound) <- c("Type", "CellID", "Alpha")

AlphaMatch <- AlphaBound[AlphaBound$CellID %in% AlphaBoundVec, ]
AlphaMatchVec <- AlphaMatch[, "CellID"]
AlphaLowland <- AlphaBound[!AlphaBound$CellID %in% AlphaMatchVec, ]

AlphaLowland$Type <- "Lowland"

AlphaLowlandVec <- AlphaLowland$CellID
AlphaLowland <- merge(AlphaLowland, LongLatDF)

AlphaLowlandScatterplot <- ggplot(AlphaLowland, aes(Latitude, Alpha)) + geom_point(shape = 16, size = 5, show.legend = FALSE, alpha=0.5, color = "cyan4") + ylab("Lowland α diversity") + ylim(0, 800) + xlab("Latitude") + theme_minimal() + theme(axis.title.y = element_text(size=32), axis.title.x = element_text(size=32),  axis.text = element_text(size=20))
AlphaLowlandScatterplot

png("Figures/AlphaMountainLowlandScatter.png", width = 1000, height = 1000, pointsize = 20)
grid.arrange(AlphaMountScatterplot, AlphaLowlandScatterplot, ncol=1)
dev.off()
```


#Bin lowlands by 5 degrees
```{r}
AlphaLowland$Bin5 = cut(AlphaLowland$Latitude,c(-65,-60,-55,-50,-45,-40,-35,-30,-25,-20,-15,-10,-5,0,5,10,15,20,25,30,35,40,45,50,55,60), labels = c("2L", "3L", "4L", "5L", "6L", "7L", "8L", "9L", "10L", "11L", "12L", "13L", "14L", "15L", "16L", "17L", "18L", "19L", "20L", "21L", "22L", "23L", "24L", "25L", "26L"))
table(AlphaLowland$Bin5)
```


#Alpha boxplot
```{r}
FullAlpha <- rbind(AlphaMount, AlphaLowland)

reordered5 <- fct_relevel(FullAlpha$Bin5, "1M","2L", "2M", "3L", "3M", "4L", "4M", "5L", "5M", "6L", "6M", "7L", "7M", "8L", "8M","9L", "9M", "10L", "10M", "11L", "11M", "12L", "12M", "13L", "13M", "14L", "14M","15L", "15M", "16L", "16M", "17L", "17M", "18L", "18M", "19L", "19M", "20L", "20M", "21L", "21M","22L", "22M", "23L", "23M", "24L", "24M", "25L", "25M", "26L")

AlphaBoxPlot <- ggplot(data = FullAlpha, aes(reordered5, Alpha, fill=Type)) + 
  geom_boxplot() + 
  ylab("α diversity") + ylim(0, 800) + xlab("Latitude") + 
  theme_minimal() + theme(axis.title.y = element_text(size=40), axis.title.x = element_text(size=40),  
                          axis.text.x = element_blank(), axis.text.y = element_text(size=20), legend.text = element_text(size=32), 
                          legend.position = "bottom", legend.title = element_blank()) + 
  scale_fill_manual(values = c("cyan4", "goldenrod2"))
AlphaBoxPlot

png("Figures/AlphaBoxPlot.png", width = 2000, height = 1000, pointsize = 20)
AlphaBoxPlot
dev.off()
```

#Alpha scatterplots combined
```{r, warning = FALSE}
FullAlpha <- rbind(AlphaMount, AlphaLowland)

AlphaScatter <- ggplot(data = FullAlpha, aes(Latitude, Alpha, color=Type)) + 
  geom_point(shape = 16, size = 3, show.legend=FALSE, alpha=0.8) + 
  ylab("α diversity") + ylim(0, 800) +
  theme_minimal() + theme(axis.title.y = element_text(size=40), axis.title.x = element_blank(),  
                          axis.text = element_text(size=20)) + 
  scale_color_manual(values = c("cyan4", "goldenrod2")) + geom_smooth(size = 2, show.legend = FALSE)
AlphaScatter
```

#SAVE ALPHA AND BETA SCATTERPLOTS :)
```{r}
png("Figures/FullScatter.png", width = 1250, height = 1000, pointsize = 20)
grid.arrange(AlphaScatter, BetaScatter, ncol=1)
dev.off()
```
