---
title: "Bryophyte Diversity Patterns Across Biomes in North and South America"
author: "Hailey Napier and Kathryn Dawdy"
output:
  html_document:
    df_print: paged
---

```{r, include = FALSE}
require(knitr)
require(latexpdf)
require(sp)
require(vegan)
require(raster)
require(rasterVis)
require(ggplot2)
require(gridExtra)
require(sf)
require(rgdal)
require(rprojroot)
require(tmap)
require(circlize)
```

```{r, setup, include = FALSE}
root <- find_rstudio_root_file()
opts_knit$set(root.dir = root)
```

```{r, include = FALSE}
#Load plot data
#Run DataProcessing2020 for plot data
BiomeBetaCellsCentCov <- readRDS("Data/BiomeBetaCellsCentCov.rds")
biomes_shp <- shapefile("Data/Biomes/Biomes_olson_projected.shp")
BiomeRange <- readRDS("Data/BiomeRange.rds")
OBRAbove100DF <- readRDS("Data/OBRAbove100DF.rds")
OBR25to100DF <- readRDS("Data/OBR25to100DF.rds")
OBR10to25DF <- readRDS("Data/OBR10to25DF.rds")
OBRBelow10DF <- readRDS("Data/OBRBelow10DF.rds")
CircleMatAll <- readRDS("Data/CircleMatAll.rds")
CircleMatAllMoss <- readRDS("Data/CircleMatAllMoss.rds")
```

```{r, include = FALSE}
#Load plot colors
cols7 <- c("#D8B70A", "#972D15", "#A2A475", "#81A88D", "#02401B",
           "#446455", "#FDD262", "#D3DDDC", "#C7B19C", "#798E87",
           "#C27D38")
biome_cols_11 <- c("#D8B70A", "#972D15", "#A2A475", "#81A88D",
                   "#02401B","#446455", "#FDD262", "#D3DDDC",
                   "#C7B19C", "#798E87", "#C27D38")

biome_cols_22 <- c(biome_cols_11, biome_cols_11)

biome_cols_87 <- c(biome_cols_22, biome_cols_11,
                   "#D8B70A", "#972D15", "#A2A475", "#81A88D",
                   "#02401B", "#446455", "#FDD262", "#D3DDDC",
                   "#C7B19C", "#C27D38", biome_cols_22, biome_cols_22)

biome_cols_66 <- c(biome_cols_22, biome_cols_22, biome_cols_22)

biome_cols_166 <- c(biome_cols_22,
                    c("#D8B70A", "#972D15", "#A2A475", "#81A88D",  "#02401B", "#FDD262", "#D3DDDC", "#C7B19C",  "#C27D38"), biome_cols_11, c("#D8B70A", "#972D15", "#A2A475", "#81A88D","#02401B", "#446455", "#FDD262", "#D3DDDC","#C27D38"), c("#D8B70A", "#972D15", "#81A88D", "#02401B", "#446455", "#FDD262", "#D3DDDC", "#C7B19C", "#798E87", "#C27D38"), c("#81A88D","#FDD262", "#D3DDDC"), c("#D8B70A", "#972D15", "#81A88D", "#446455", "#D3DDDC", "#C7B19C", "#798E87"),
c("#D8B70A", "#972D15", "#A2A475", "#81A88D", "#02401B","#FDD262", "#D3DDDC", "#C7B19C", "#C27D38"),"#81A88D", c("#972D15", "#A2A475", "#81A88D", "#02401B", "#FDD262", "#D3DDDC", "#C7B19C", "#C27D38"),
c("#D8B70A","#446455", "#FDD262", "#D3DDDC"), biome_cols_22, c("#D8B70A", "#972D15", "#81A88D","#D3DDDC", "#C7B19C", "#C27D38"), c("#D8B70A", "#446455", "#FDD262", "#D3DDDC", "#798E87", "#C27D38"), c("#A2A475", "#81A88D", "#02401B","#FDD262", "#D3DDDC", "#C7B19C"),
c("#D8B70A", "#972D15", "#A2A475", "#81A88D",  "#02401B", "#FDD262", "#D3DDDC", "#C7B19C", "#C27D38"),biome_cols_11, c("#D8B70A", "#C27D38"), biome_cols_11)

biome_cols_29 <- c(biome_cols_11,
                   c("#D8B70A", "#972D15", "#A2A475", "#81A88D","#02401B", "#446455", "#FDD262", "#D3DDDC", "#798E87","#C27D38"), c("#972D15", "#A2A475", "#81A88D", "#02401B",
"#FDD262", "#D3DDDC", "#C7B19C", "#C27D38"))

grid.col <- c(Coniferous_Forests = "#D8B70A", Dry_Forest = "#972D15", 
              Mediterranean_Woodlands = "#A2A475", Moist_Forest = "#81A88D", 
              Savannas = "#02401B", Taiga = "#446455", Temperate_Grasslands = "#FDD262", 
              Temperate_Mixed = "#D3DDDC", Tropical_Grasslands = "#C7B19C", Tundra = "#798E87", 
              Xeric_Woodlands = "#C27D38", Hypnales = "grey", Porellales = "grey",  
              Pottiales = "grey", Hookeriales  = "grey", Bryales = "grey", Jungermanniales = "grey", 
              Andreaeaeales = "grey", Splachnales = "grey", Orthotrichales  =  "grey",  
              Bartramiales = "grey", Metzgeriales = "grey", Dicranales = "grey", Anthocerotales = "grey",
              Funariales = "grey", Treubiales = "grey", Archidiales = "grey", Marchantiales = "grey", 
              Polytrichales = "grey", Aulacomniales = "grey", Fossombroniales = "grey", 
              Grimmiales = "grey", Hedwigiales = "grey", Bryoxiphiales = "grey", Buxbaumiales  = "grey",
              Ptychomniales = "grey", Gigaspermales = "grey", Dendrocerotales  = "grey", 
              Pleuroziales = "grey", Rhizogoniales = "grey", Haplomitriales = "grey", Hypnodendrales  = "grey", 
              Pallaviciniales = "grey",  Pelliales = "grey", Notothyladales = "grey", Ricciales = "grey", 
              Ptilidiales = "grey", Sphaerocarpales = "grey", Sphagnales = "grey")
```


### Biomes in North and South America
```{r my-chunk, fig.height=6, fig.width=10, fig.align="center"}
BiomeMap <- qtm(biomes_shp,
                        fill="biomes", 
                        fill.style="fixed",
                        fill.labels=biomes_shp$biomes,
                        fill.palette=cols7,
                        fill.title="Biomes",
                        layout.legend.position=c("left","bottom"),
                        layout.legend.width=1.5,
                        layout.frame=FALSE)
BiomeMap
```
<center>
*Figure 1. Map of North and South American biomes.* 
</center>
  
     
    
      
### Beta Diversity Plots by Biome
```{r fig.height=9, fig.width=15, fig.align="center", warning=FALSE}
BiomeBetaWeightBV <- ggplot(BiomeBetaCellsCentCov, aes(x=Type, y=Beta, fill=Type, color=Type)) + 
  geom_boxplot(show.legend = FALSE, fill=biome_cols_11, color="black") +
  guides(x = guide_axis(angle=30)) +
  #theme_minimal() +        #un-comment whichever theme you want
  theme_gray() +
  #theme_light() +
  #theme_bw() +
  geom_violin(scale="count", show.legend=FALSE, fill="gray", alpha=0.35,
              color="gray25") +
  xlab("Biome") +
  ylab("Beta Diversity") +  
  theme(axis.title.y = element_text(size=22), 
        axis.title.x = element_text(size=22),
        axis.text.y = element_text(size=16), 
        axis.text.x = element_text(angle = 30, hjust = 1, size = 16))
BiomeBetaWeightBV
```
  
  
    
      
### Range Size Plots by Biome
```{r fig.height=9, fig.width=15, fig.align="center", warning=FALSE}
BiomeRangeBV <- ggplot(BiomeRange, aes(x=Type, y=Range, fill=Type, color=Type)) + 
  geom_boxplot(show.legend = FALSE, fill=biome_cols_11, color="black") +
  guides(x = guide_axis(angle=30)) +
  #un-comment whichever theme you want
  #theme_minimal() +        
  theme_gray() +
  #theme_light() +
  #theme_bw() +
  geom_violin(scale="count", show.legend=FALSE, fill="gray", alpha=0.35,
              color="gray25") +
  xlab("Biome") +
  ylab("Range") +  
  theme(axis.title.y = element_text(size=22), 
        axis.title.x = element_text(size=22),
        axis.text.y = element_text(size=20), 
        axis.text.x = element_text(angle = 30, hjust = 1, size = 16))
BiomeRangeBV
```
  
    
      
        
### Alpha Diversity by Order, Split into Biomes
#### Orders with Maximum Alpha Diversity < 10
```{r fig.height=9, fig.width=15, fig.align="center", warning=FALSE}
FacetOBRBelow10 <- ggplot(OBRBelow10DF, aes(x=Biome, y=Alpha, fill=Biome, color=Biome)) + 
  geom_boxplot(show.legend = FALSE, fill=biome_cols_166, color="black", outlier.size=0.7) +
  #un-comment whichever theme you want
  #theme_minimal() +
  theme_gray() +
  #theme_light() +
  #theme_bw() +
  #theme_dark() +
  #theme_linedraw() +
  #geom_violin(scale="count", show.legend=FALSE, fill="gray", alpha=0.35, color="gray25") +
  xlab("Biome") +
  ylab("Richness") +  
  theme(axis.title.y = element_text(size=22), 
        axis.title.x = element_text(size=22),
        axis.text.y = element_text(size=15), 
        axis.text.x = element_text(angle = 90, hjust = 1, size = 12))+
  facet_wrap(~Order,
             #un-comment # of rows you want
             ncol=6 
             #ncol=5
             )
FacetOBRBelow10
```
   
      
#### Orders with Maximum Alpha Diversity between 10 and 25
```{r fig.height=9, fig.width=17, fig.align="center", warning=FALSE}
FacetOBR10to25 <- ggplot(OBR10to25DF, aes(x=Biome, y=Alpha, fill=Biome, color=Biome)) + 
  geom_boxplot(show.legend = FALSE, fill=biome_cols_66, color="black", outlier.size=0.7) +
  #un-comment whichever theme you want
  #theme_minimal() +
  theme_gray() +
  #theme_light() +
  #theme_bw() +
  #theme_dark() +
  #theme_linedraw() +
  geom_violin(scale="count", show.legend=FALSE, fill="gray", alpha=0.35, color="gray25") +
  xlab("Biome") +
  ylab("Richness") +  
  theme(axis.title.y = element_text(size=22), 
        axis.title.x = element_text(size=22),
        axis.text.y = element_text(size=15), 
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12))+
  facet_wrap(~Order,
             #un-comment # of rows you want
             ncol=3
             #ncol=2
             )
FacetOBR10to25
```
  
    
#### Orders with Maximum Alpha Diversity between 25 and 100
```{r fig.height=10, fig.width=16, fig.align="center", warning=FALSE}
FacetOBR25to100 <- ggplot(OBR25to100DF, aes(x=Biome, y=Alpha, fill=Biome, color=Biome)) + 
  geom_boxplot(show.legend = FALSE, fill=biome_cols_87, color="black", outlier.size=0.7) +
  #un-comment whichever theme you want
  #theme_minimal() +
  theme_gray() +
  #theme_light() +
  #theme_bw() +
  #theme_dark() +
  #theme_linedraw() +
  geom_violin(scale="count", show.legend=FALSE, fill="gray", alpha=0.35, color="gray25") +
  xlab("Biome") +
  ylab("Richness") +  
  theme(axis.title.y = element_text(size=22), 
        axis.title.x = element_text(size=22),axis.text.y = element_text(size=15), 
        axis.text.x = element_text(angle = 30, hjust = 1, size = 10)) +
  facet_wrap(~Order, 
             #un-comment # of rows you want
             ncol=4
             #ncol=2
             )
FacetOBR25to100
```
  
    
#### Orders with Maximum Alpha Diversity > 100
```{r fig.height=9, fig.width=15, fig.align="center", warning=FALSE}
FacetOBRAbove100 <- ggplot(OBRAbove100DF, aes(x=Biome, y=Alpha, fill=Biome, color=Biome)) + 
  geom_boxplot(show.legend = FALSE, fill=biome_cols_22, color="black",  outlier.size=1) +
  #un-comment whichever theme you want
  #theme_minimal() +
  theme_gray() +
  #theme_light() +
  #theme_bw() +
  #theme_dark() +
  #theme_linedraw() +
  geom_violin(scale="count", show.legend=FALSE, fill="gray", alpha=0.35, color="gray25") +
  xlab("Biome") +
  ylab("Richness") +  
  theme(axis.title.y = element_text(size=22), 
        axis.title.x = element_text(size=22),
        axis.text.y = element_text(size=15), 
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12)) +
  facet_wrap(~Order, 
             #un-comment # of rows you want
             ncol=2
             #ncol=1
             )
FacetOBRAbove100
```
  
  
    
       
### Circle Plots 
#### *Species richness for each order, grouped by biome*
#### All bryophyte orders
```{r fig.height=9, fig.width=9, fig.align="center"}
circos.clear()
circos.par(start.degree = 0)
chordDiagram(CircleMatAll, grid.col = grid.col, column.col = biome_cols_11, 
             directional = 1, direction.type = "arrows", link.arr.type = "big.arrow", 
             link.arr.length = 0.05, link.largest.ontop = T, annotationTrack = c("grid"), 
             preAllocateTracks = 1, big.gap = 20, small.gap = 2)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  xlim = get.cell.meta.data("xlim")
  ylim = get.cell.meta.data("ylim")
  sector.name = get.cell.meta.data("sector.index")
  circos.text(mean(xlim), ylim[1], sector.name, facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5), cex=0.5)
}, bg.border = NA)
```
  
    
#### Only moss orders
```{r fig.height=9, fig.width=9, fig.align="center"}
circos.clear()
circos.par(start.degree = 0)
chordDiagram(CircleMatAllMoss, grid.col = grid.col, column.col = biome_cols_11, 
             directional = 1, direction.type = "arrows", link.arr.type = "big.arrow", 
             link.arr.length = 0.05, link.largest.ontop = T, annotationTrack = c("grid"), 
             preAllocateTracks = 1, big.gap = 20, small.gap = 2)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  xlim = get.cell.meta.data("xlim")
  ylim = get.cell.meta.data("ylim")
  sector.name = get.cell.meta.data("sector.index")
  circos.text(mean(xlim), ylim[1], sector.name, facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5), cex=0.5)
}, bg.border = NA)
```

